---
title: "Getting And Cleaning Data"
author: "CX"
date: "07/19/2014"
output: html_document
---
Data source: https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

I put the unzipped data in the following directory
'/Users/xc/datasciencecoursera/GetClean_Assign1/UCI HAR Dataset'

If you plan to run the script, you might need to change the working directory.

The script will produce dataframes X, Y, which are required by steps 1-4 in the assignment, and a file 'ave_data.txt' in the working directory, whose content is the tidy dataset required by step 5.

Set working directory, this is the line that needs to be changed if you plan to run the script on your computer.
```{r}
setwd('/Users/xc/datasciencecoursera/GetClean_Assign1')
```

Read feature names
```{r}
feature_names <- read.table('./UCI HAR Dataset/features.txt', 
                            comment.char = '', nrows = 561, 
                            col.names = c('temp', 'names'))
feature_names[, 1] <- NULL
head(feature_names)
```

Read training data, giving feature names as column names
```{r}
subject_train <- 
  read.table('./UCI HAR Dataset/train/subject_train.txt')
m_train <- length(subject_train[, 1])
X_train <- read.table('./UCI HAR Dataset/train/X_train.txt',
                      comment.char = '', colClasses = 
                        rep('numeric', 561), col.names = 
                        feature_names[, 1], nrows = m_train)
Y_train <- read.table('./UCI HAR Dataset/train/Y_train.txt', 
                      comment.char = '', colClasses = 'numeric',
                      nrows = m_train, col.names = 'labels')
head(names(X_train))
```

Read testing data, does the exactly same thing as above with testing data
```{r}
subject_test <- 
  read.table('./UCI HAR Dataset/test/subject_test.txt')
m_test <- length(subject_test[, 1])
X_test <- read.table('./UCI HAR Dataset/test/X_test.txt',
                      comment.char = '', colClasses = 
                        rep('numeric', 561), col.names = 
                        feature_names[, 1], nrows = m_test)
Y_test <- read.table('./UCI HAR Dataset/test/Y_test.txt', 
                      comment.char = '', colClasses = 'numeric',
                      nrows = m_test, col.names = 'labels')
```

Merge traing and testing data, X_temp, Y_temp is going to be 
replaced by X and Y in the next step
```{r}
X_temp <- rbind(X_train, X_test)
Y_temp <- rbind(Y_train, Y_test)
```

Process the subject variable corresponding to our data, which will be appended to X and Y later on
```{r}
subject <- rbind(subject_train, subject_test)
names(subject) <- 'subject'
subject <- transform(subject, subject = as.factor(subject))
head(subject, 3)
tail(subject, 3)
```

Extract only the mean and std for each measurements using regular expressions, then merge the new data set. I excluded variables with names like 'meanFrenquency' here.
```{r}
meanX <- X_temp[, grep('.*mean()[^F]', names(X_temp))]
stdX <- X_temp[, grep('.*std()', names(X_temp))]
X <- cbind(meanX, stdX)
```

Read in activity names and assign it to Y
```{r}
act_names <- read.table('./UCI HAR Dataset/activity_labels.txt')
temp <- as.factor(Y_temp$labels)
levels(temp) <- act_names$V2
Y <- transform(Y_temp, labels = temp)
```

Finally, merge subject and X, subject and Y, and this will be the dataframe requested by step 1-4
```{r}
X <- cbind(subject, X)
Y <- cbind(subject, Y)
head(Y, 3)
tail(Y, 3)
```

Calculate the tidy data set using dcast. Here I do this on the new data set with only mean and std extracted from the original big data set, since I can't tell exactly on which data set I shoud work on from the instructions. To do this on the oringal data set, simply apply the following steps to the dataframes X_temp calculated previously(with Y added).
```{r}
library(plyr)
library(reshape2)
my_data <- cbind(X, 'labels' = Y$labels)
my_factor <- with(my_data, interaction(subject, labels))
my_data <- melt(my_data, id = c('subject', 'labels'),
                 measure.vars = names(my_data)[2:67])
my_data <- dcast(my_data, subject + labels ~ variable, mean)
write.table(my_data, file = 'ave_data.txt', )
```
And ave_data.txt will be the tidy data set that gets uploaded.